Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-02 paper-2024-sac-marl-neighbouring-strategies.tex
chore: add the pseudocode

\usepackage{booktabs} % For formal tables
\usepackage{graphicx}
\usepackage{autobreak}
\usepackage{mytodonotes}
%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{cleveref}

@@

% \author{Nicolò Malucelli}
% \orcid{1234-5678-9012}
% \affiliation{%
%   \institution{University of Bologna}
%   \streetaddress{Via dell'Università, 50}
%   \city{Cesena} 
%   \country{Italy}
% }
% \email{nicolo.malucelli@studio.unibo.it}

% \author{Davide Domini}
% \orcid{1234-5678-9012}
% \affiliation{%
%   \institution{University of Bologna}
%   \streetaddress{Via dell'Università, 50}
%   \city{Cesena} 
%   \country{Italy}
% }
% \email{davide.domini@unibo.it}

% \author{Gianluca Aguzzi}
% \affiliation{%
%   \institution{University of Bologna}
%   \streetaddress{Via dell'Università, 50}
%   \city{Cesena} 
%   \country{Italy}
% }
% \email{gianluca.aguzzi@unibo.it}

% \author{Mirko Viroli}
% \affiliation{%
%   \institution{University of Bologna}
%   \streetaddress{Via dell'Università, 50}
%   \city{Cesena} 
%   \country{Italy}
% }
% \email{mirko.viroli@unibo.it}

% % The default list of authors is too long for headers}
% \renewcommand{\shortauthors}{N. Malucelli et al.}

@@

\subsection{Nearest Neighboring Consensus}

\subsection{Experience Sharing}

\begin{itemize}
\item Provide a detailed explanation of each neighbor-based method:
\begin{itemize}
\item \textbf{Independent learners:} Emphasize its simplicity and scalability but highlight its vulnerability to non-stationarity.
\item \textbf{NN-Averaging:} Describe the weight averaging process, discuss the influence of neighbor selection and averaging frequency, and mention weighted averaging variations.
\item \textbf{NN-Consensus:} Explain the selection of the best-performing network and its potential benefits and drawbacks.
\item \textbf{Experience Sharing:} Detail how agents share experiences, discuss asynchronous vs. synchronous sharing, and highlight its advantages in sparse interaction environments.
\end{itemize}
\end{itemize}
\section{Experimental Setup}\label{sec:experiments}
\begin{itemize}
\item Introduce the custom "collect the items" multi-agent environment:
\begin{itemize}
\item Describe the environment's dynamics, agent goals, and challenges.
\item Specify the action space (continuous and discrete) and observation space design.
\item Explain the reward structure, emphasizing the components that encourage cooperation.
\end{itemize}
\item Detail the implementation framework (Gymnasium, RLlib, PyTorch).
\item Specify the chosen RL algorithm (DQN) and its hyperparameter configuration.
\item Define the evaluation metrics: average episode length and average reward.
\end{itemize}

>>>>

\usepackage{booktabs} % For formal tables
\usepackage{graphicx}
\usepackage{autobreak}
\usepackage{mytodonotes}
\usepackage[ruled,vlined]{algorithm2e}
%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{cleveref}

@@

%   \institution{University of Bologna}
%   \streetaddress{Via dell'Università, 50}
%   \city{Cesena} 
%   \country{Italy}
% }
% \email{davide.domini@unibo.it}

% \author{Gianluca Aguzzi}
% \affiliation{%
%   \institution{University of Bologna}
%   \streetaddress{Via dell'Università, 50}
%   \city{Cesena} 
%   \country{Italy}\usepackage{algorithm}

% \renewcommand{\shortauthors}{N. Malucelli et al.}

@@

\subsection{K-Nearest Neighboring Consensus}
In case of homogeneous agents (i.e., when all agents have the same action and observation space),
the goal of learning is to find a common policy that all agents can follow.
This approach is typically used in large-scale systems where agents are identical and need to reach a consensus---see [cite].
%
In this case, instead of forcing a common policy, we can let agents learn independently and then select the best-performing agent as the reference for the others, leading to a consensus on the best policy.
%
In this way, each agent $i$ continue to learn independently, but at each iteration, it updates its Q-network using the Q-values of the best-performing agent in its neighborhood.

The best performing agent is selected based on the average reward it has received in the last $n$ episodes.
Formally, the update rule for agent $i$ at time $t+1$ is:
\begin{equation}
  \theta^{t+1}_i \leftarrow \theta^t_{\text{best}}
\end{equation}
where $\theta^t_{\text{best}}$ is the Q-network of the best-performing agent in the neighborhood $\mathcal{N}_k(i)$ at time $t$.

\subsection{Experience Sharing}
Experience sharing is a simple yet effective method for distributed learning.
This approach is typically used in centralized training to improve sample efficiency and learning stability.
In our distributed setting, each agent maintains its own experience replay buffer, but it also shares a fraction of its experiences with its neighbors.
This allows agents to learn from each other's experiences, potentially accelerating learning and improving coordination because agents can learn from each other's mistakes and successes.
%
Formally, the reply buffer of agent $i$ at time $t$ is $\mathcal{D}^t_i = \{(o^i, a^i, r^i, o'^i)\}$.
At each iteration, agent $i$ shares a fraction $\beta$ of its experiences with its neighbors in $\mathcal{N}_k(i)$.
The update rule for agent $i$ at time $t+1$ is:
\begin{equation}
  \mathcal{D}^{t+1}_i = \mathcal{D}^t_i \cup \bigcup_{j \in \mathcal{N}_k(i)} \beta \mathcal{D}^t_j
\end{equation}
\subsection{Neighboring-Based Training with Deep Q-Learning}
The neighboring-based training strategies can be integrated with deep Q-learning to enable distributed learning in multi-agent settings.
%
The algorithm follow the same structure as the deep Q-learning algorithm, but with the addition of the neighboring-based intereactions.
%
This may happen with a different frequency for each strategy, depending on the communication overhead and the desired level of coordination.

In [ref] we provide a detailed description of the integration of the neighboring-based strategies with deep Q-learning.
\begin{algorithm}
  \caption{Neighboring-Based Deep Q-Learning}
  \label{alg:neighbor_dqn}
  \ForEach{agent $i \in N$}{
      Initialize Q-network $Q^i(o^i, a^i; \theta^i)$ and target network $Q^{i-}(o^i, a^i; \theta^{i-})$\\
      Initialize replay buffer $\mathcal{B}^i$
  }
  \For{episode = 1 to M}{
      \ForEach{agent $i \in N$}{
          Initialize initial observation $o_0^i$
      }
      \For{t = 0 to T}{
          \ForEach{agent $i \in N$}{
              Select action $a_t^i$ based on $\epsilon$-greedy policy using $Q^i(o_t^i, a^i; \theta^i)$\\
              Execute action $a_t^i$ and observe reward $r_{t+1}^i$ and next observation $o_{t+1}^i$\\
              Store $(o_t^i, a_t^i, r_{t+1}^i, o_{t+1}^i)$ in $\mathcal{B}^i$\\
              Sample a mini-batch from $\mathcal{B}^i$\\
              Calculate target values using chosen neighboring strategy (see below)\\
              Update $\theta^i$ by minimizing the TD error\\
              Every C steps, update target network $\theta^{i-} \leftarrow \theta^i$\\
              \textbf{
              Every S steps, perform neighbor interaction (e.g., averaging, consensus, experience sharing)}
          }
      }
  }
\end{algorithm}

\section{Experimental Setup}\label{sec:experiments}
\begin{itemize}
\item Introduce the custom "collect the items" multi-agent environment:
\begin{itemize}
\item Describe the environment's dynamics, agent goals, and challenges.
\item Specify the action space (continuous and discrete) and observation space design.
\item Explain the reward structure, emphasizing the components that encourage cooperation.
\end{itemize}
\item Detail the implementation framework (Gymnasium, RLlib, PyTorch).
\item Specify the chosen RL algorithm (DQN) and its hyperparameter configuration.
\item Define the evaluation metrics: average episode length and average reward.
\end{itemize}
